Alleviating the Topology Mismatch Problem in Distributed Overlay Networks

### 2014-01-30 structured.tex
2013.05.26 (Vassilis) Structured algorithms section reworked.

\section{Structured P2P Networks}
\label{section:structured}

This section presents a description and analysis of algorithms proposed
to address the topology mismatch problem in structured P2P networks.
In structured P2P networks, the construction of the routing tables
held by participating nodes determines the efficiency of message forwarding.
Routing tables that closely match the underlying IP topology
achieve increased performance. As a result, the proposed solutions for
the topology mismatch problem for structured P2P networks 
use different levels of proximity
information to optimize the routing tables used by peers.  
The categorization of the algorithms we present here is based on the
categorization suggested by previous 
work~\cite{CDHR2002,CDCR2002,RSS2002}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Global Soft-State}

@@

\textit{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network to the overlay. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state is made available to all nodes to help them find the best way to
route their messages. \textit{Global Soft-State} operates in two phases: 
generation and use of the proximity information. For generating the proximity
information a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm chooses a different path from the
classic gossiping approaches for constructing and maintaining the overlay. It is
based on landmark clustering for strategic placement of proximity information on
the overlay enabling any node to access such information using a landmark number
that reflects its physical position in the network. For various logical
regions\footnote{This might be a high-order zone in the eCAN\cite{xu_ecan_2002}
context or a set of nodes sharing a particular prefix in overlays such as
Pastry\cite{antony_pastry_2001}.} maps of physical information are built and
published where each node may appear in a maximum of $log\left( N \right)$ such
maps. To dynamically adapt to changing network conditions, a node subscribes to
relevant \emph{soft states} that utilize a notification system in order to
initialize any necessary neighbor re-selection.

Maintainece of several host states at different layers, renders any content
migration a costly process. Additionally, the method does not make any
continuing effort to remap
the overlay structure after nodes successfully join resulting into poor
adaptation to changing conditions. Although this approach greatly reduces the
routing latency to far nodes, it is unable to dynamically identify nodes that
are close to routers and gateways in order to construct the secondary overlay.
Nevertheless, static recognition of such nodes is currently done based on BGP
reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
traditional DHTs.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Unable to identify nodes that are close to routers but it can reduce routing
% latency to far away node by approximately 50-75%.
medium &
% maintenance of node states, subscriptions and notifications need global
% overview of the overlay resulting in high control overhead
high &
% Statically implemented sacrificing the decentralized nature of DHTs and
% thus poorly scaling.
low
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Mithos}
\emph{Mithos} \cite{WR2003} is a P2P protocol which incorporates directed
incremental probing to find near optimal node placement, and is an integration
of geographic layout and proximity routing overlay optimization methods.

The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
the first set of candidates, and while the iteratively closest nodes are detected by
probing the neighborhood, the candidate neighbor list is updated. To
avoid a local minima, Mithos probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

After finding its first neighbor, the newcoming peer is assigned an ID using
information gathered during the iterative neighbor selection phase. Virtual
coordinates are assigned to the newcoming peer by using the distances of the two
closest nodes and their neighbors, so that Euclidean distances between the node
and all known hosts predict the network latency between them
\cite{cox_vivaldi_2004}. The benefit of these synthetic coordinates are that
they are explicitly used as the node's ID, and distance computations among nodes
can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbors. Mithos
uses a \emph{quadrant-based} mechanism in which each node establishes
a link to the closest neighbor in each quadrant. During forwarding, the next
hop is performed towards a neighbor in the same quadrant as the final
destination. The new coming node may not know of other neighbors in all
quadrants, therefore, the node first identifies neighbors in all quadrants
using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
the Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
results using parallel path processing by taking into account further geometric
properties of node relationships.

One major limitation of \textit{Mithos} is that the protocol cannot handle
dynamic arrivals and departures from the network.

%
% TODO: review this part
%
%In order to avoid local minima during neighbor detection, extensive probing
%  must be undertaken. In simulation, unfortunately, only very small-sized
% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
% conclusions can be made as for the behavior of an extensively large,
% real-world p2p deployment of the scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{LAPTOP}
\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
with the goal of reducing hops during message routing as well as minimizing
maintenance overhead. Additionally, a caching scheme is also incorporated so as
to further reduce routing table update costs. The authors theoretically show
that in LAPTOP, routing path length is bounded by $O(log_d N)$ and node joining
and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
the maximum degree of each node. LAPTOP implements the geographical layout
approach  and constructs its layout in a self-organizing and efficient fashion.
The physical distances are estimated with round trip times to a few existing
nodes in the overlay network.

@@

The routing scheme is similar to the longest-prefix IP matching scheme. At each
forwarding hop, any message travels up the tree until the first common ancestor
of both source and destination nodes is reached and then starts descending to
arrive to its target. During tree traversals, special entries in the routing
tables, called \emph{routing cache}, are maintained to increase routing
efficiency and achieve finer load balance. Caching enables a node to forward a
message to a better longest-prefix match than that of its direct ancestor making
a larger, quicker and more cost effective step through the overlay and toward
the destination. To improve scalability, the number of child nodes and the
size of the routing cache are limited. In terms of overlay maintenance, LAPTOP
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Proximity in Kademlia}
% Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
% summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
% \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
% protocol for P2P networks, which uses an iteration based lookup algorithm. The
% protocol uses the standard $160$ bit ID system for nodes and locates the nodes
% in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
% operations are done over this prefix binary tree, which converges to
%logarithmic
% lookup times. The ID's are assigned randomly, therefore in Kademlia, there is
%no
% proximity control, which results in inefficient use of the underlying network
% during lookup and retrieve operations.

\emph{Proximity in Kademlia} \cite{KLKP2008} introduces proximity
controls over the vanilla Kademlia protocol's random ID assignment to
optimize the underlying network usage of the protocol. The work in this paper
can also be considered as a more generic study of incorporating proximity into
iterative lookup algorithms. 
The overlay tier improves its routing
performance using a cost metric that is provided by the underlying tier
and quantifies the suitability of establishing a communication link between peers as
a cost function, based on the used proximity criteria (RTT, ISP locality, etc.).
%An underlay metric provides information about the underlay network. The paper
%  distinguishes between three kinds of possibilities to quantify it:
%\begin{itemize}
% \item Using measurements gained by previous lookups
% \item Using measurements acquired by other peers or jointly calculated with
%  others
% \item Using a local database to look up information
%\end{itemize}

The paper proposes two approaches for overlay optimization: 
{\em proximity routing} and {\em proximity neighbor selection}.
In proximity routing, the goal
is to choose the best next hop during routing a message. Due
to the fact that the routing in Kademlia is iterative, it is the initiator of
the lookup that chooses each next hop from a set of candidate nodes. Standard
Kademlia always chooses the closest node with respect to the XOR metric but with
proximity routing selection, Kademlia chooses the node with the smallest underlay
metric cost. Thus, as in all such approaches, there is a trade-off between the
overlay and underlay distances.

Towards the enhancement of traffic locality, the authors also use information
from the MaxMind GeoIP database\footnote{MaxMind Geolocation Technology.
http://www.maxmind.com} to retrieve proximity information of a given peer based
on its region, country, or ISP location, and use this information to form
clusters of nearby peers. An alternative method could use Vivaldi
coordinates \cite{cox_vivaldi_2004}. However, the authors report that their clustering
approach performs better than the Vivaldi protocol based on their experiments.
Authors also report that {\em Proximity routing} protocol successfully worked
with the Kademlia protocol and improved the locality of the connections between
peers.

In \emph{proximity neighbor selection}, the goal is to
keep peers with the least contact cost in the routing table. As Kademlia
continuously learns about new peers via incoming requests and an iterative
lookup process, no 
special algorithm for searching more cost-effective peers is necessary and
it simply chooses the best peers seen so far.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Uses static information to cluster nodes which is not the most efficient
% approach in dynamic systems
medium &
% 
medium &
%
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
\emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
Chord\cite{stoica_chord_2001}, CHOP6 uses a finger table, whose entries hold
more than one candidate node.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Simple approach that exploits proximity but it is coarse grained and thus not
% efficient.
low &
%
low &
%
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{T2MC}

T2MC \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
cluster nearby peers together to reduce the redundant multiple message
passes between the AS boundaries. T2MC uses a customized k-mean classification
algorithm with k = 2 to perform the classification, and exploits the stable
structure of the internet routers to guide clustering. Even though traceroute
provides detailed information about the network structure, use of traceroute
creates overhead on the overall network structure. For this same reason, it is
not unusual for network administrators to disable this support on their network
routers, which may affect the performance of the T2MC algorithm.

%T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
%systems and cluster close by peers, thus reducing redundant traffic that crosses
%autonomous system boundaries. T2MC uses a customized k-mean classification
%algorithm with $k=2$ to perform the classification, and exploits the stable
%structure of the Internet routers to guide clustering. The peer chooses the
%minimum and maximum Latency results from the trace-route to initialize the
%centroids of two sets. Then the peer calculates, for all hops along its
%trace-routed path, the peer calculates the absolute distance to these centroids
%of both sets and assigns the routers to that centroid with which it has the
%smaller absolute distance. After that, the mean and variance values are
%calculated and if the variance is larger than a predefined threshold then the
%algorithm sets the two latency mean values as new centroids for the two sets and
%loops again. Ultimately peer will end up with two sets having the minimum
%intra-set variance. The peer, then, chooses the router from the second set with
%the minimum hops attribute and sets it as a threshold. The selected router and
%all others whose hops attribute is larger than the threshold are classified as
%\emph{remote} router cluster. The remaining are classified as \emph{near}. From
%the near class, the peer chooses the one with maximum hops attribute as its edge
%router, and registers it along with all near clusters into the DHT of the
%P2P overlay. As new peers join the network, those that share the same edge
%router or any of the members of the near router clusters, would form a close
%peer cluster. As Edge routers can provide more valuable information than other
%members of the near set, T2MC was designed to prioritize interaction of peers
%and edge gateways.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{PChord}
\emph{PChord}\cite{HLYW2005} is based on the Chord DHT and adds
proximity routing into its routing mechanism. The main modification over the
vanilla Chord is the inclusion of a \emph{proximity list} into its routing table
so that the next hop is decided based on both the best progress
towards the key and the physical proximity of the candidate nodes as well.
The list is, at join time, empty but as the PChord node starts to interact with
other nodes it applies a heuristic mechanism to fill the list. Entries are
dynamically added or removed as the network state changes. The overlay
maintenance cost can be considered similar to Chord's. Besides node join/leave
and cost of heartbeat messages for connection verification, PChord adds
heartbeat of nodes inside the proximity list but since this list always contains
nodes with the lowest RTT the added cost is kept to a minimum.

Due to the fact that, with each step, there is always progress towards the
target node in the ID space, PChord will result in a reduced hop number than
Chord, since each hop in PChord is larger or at least equal in the key space.
Passing through proximity links in the underlying network means reduced routing
cost. Additionally, if the number of entries in the proximity list is the same
as the number of network partitions, PChord prevents hops from jumping in and
out of the network partition the current node belongs to.

% TODO: SOME DISCUSSION
% TODO: NEEDS REWORDING
% PChord’s main routing optimizations are of less overlay hops, passing
% proximity links of the underly network and passing the same network partition
% only once in the routing path, which directly results in the lower RDP of
% routing efficiency. Meanwhile, PChord keeps light-weight maintenance cost as
% Chord does.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Small decrease in needed overlay hops per object publishing
medium &
% The are heartbeat mechanism for both finger table and proximity list adds a
% burden to the network by default. But the overall overhead is decreased as
% the proximity list becomes longer.
medium &
% Improvement of routing is achieved with more object pointers helping to
% shorten the routing path of object querying. When the distance of query source
% to target is far away both Chord and PChord achieve the average hop number to
% not increase.
high
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{AChord}
\emph{AChord} \cite{DK2006} uses IPv6's anycast mechanism to
relieve the protocol from complex joining procedures, and achieve high accuracy
of network proximity and thus, high routing efficiency. Anycast delivers a message
coming from the outside of an anycast group to the physically closest node in
that group. AChord organizes all nodes participating in the overlay network into
an anycast group. Any joining node comes outside of that group and, thus, is
automatically forwarded to the physically nearest node in order to bootstrap,
avoiding the need to, explicitly, maintain such nodes, and the effort of finding
a way of locating the physically nearest among them. The ID of the newcoming
node is computed based on the ID of the bootstrap node and the bootstrap's
predecessor in a way that the constructed ID will be positioned between the
aforementioned two. After joining, the finger table is built the same way as in
the Chord protocol. Moreover, additional nodes are maintained in a structure
called \emph{adjacency table} which stores information about the closest known
node. The routing decision is made by using both the adjacency and the finger
table. The authors claim this to be a lightweight approach that requires few
changes to Chord while keeping its advantageous characteristics.

>>>>

\section{Structured \p\ Networks}
\label{section:structured}

This section presents a description and analysis of algorithms proposed
to address the topology mismatch problem in structured \p\ networks. The
categorization of the approaches we present here is based on how different
levels of proximity information is gathered and the way peer routing tables are
maintained and used in an effort make overlay hops as efficient as possible. The
categorization is based on previous work found in the
literature~\cite{CDHR2002,CDCR2002,RSS2002}.

%% VM Rephrased as above to keep it analogous to the opening paragraph of
%% the unstructured part.
%%
% This section presents a description and analysis of algorithms proposed
% to address the topology mismatch problem in structured \p\ networks.
% In structured \p\ networks, the construction of the routing tables
% held by participating nodes determines the efficiency of message forwarding.
% Routing tables that closely match the underlying IP topology
% achieve increased performance. As a result, the proposed solutions for
% the topology mismatch problem for structured \p\ networks
% use different levels of proximity
% information to optimize the routing tables used by peers.
% The categorization of the algorithms we present here is based on the
% categorization suggested by previous
% work~\cite{CDHR2002,CDCR2002,RSS2002}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Global Soft-State}

@@

\emph{Global Soft-State} \cite{XTZ2003} builds a global map to help
choose shorter routing paths, combining the landmark binning method and small
scale distance probes to reveal the proximity properties of the underlying
network. %Figure \ref{fig:global_softstate} presents the landmark
%space and the mapping of the position of the nodes.
This global view of the
state, is made available to all nodes to help them find the best way to
route their messages. \emph{Global Soft-State} operates in two phases:
generation and use of the proximity information. For generating the proximity
information, a hybrid approach is proposed, which uses landmark clustering as a
preprocessing step to select a number of potential nearest neighbor
candidates and then refine the selection by incorporating a
round-trip-time-based scheme to ultimately choose the closest node. For using
the proximity information, the algorithm follows a different path from that of the
classic gossiping approaches for the construction and maintenance of the
overlay. It is based on landmark clustering for generation and strategic
placement of proximity information on the overlay. This information is called a
``map'' and concerns a region of the overlay (part of the Cartesian space in
structures like eCAN\cite{xu_ecan_2002} or a set of nodes sharing a particular
prefix in overlays such as Pastry\cite{antony_pastry_2001}). Thus, each node not
only can access such information from these landmarks but can also subscribe to
relevant \emph{soft states} and get notified on dynamic network changes. A
node may appear in a maximum of $log\left( N \right)$ such
maps.
%%
Although this approach greatly reduces the routing latency to far nodes, it
comes expensive. It is possible to end up with a very large number of regions
each of which has to maintain a map. The fact that a peer may appear in multiple
such maps we can conclude that additional traffic for probing and notifications
renders the approach does not scalable. Moreover, to refine measurements
additional inner-bin measurement s are required\cite{WZS2004}. Also improvement
is done within a region because of little knowledge about neighboring zones.

%% VM Removed most of the paragraph below which contains info that we cannot back
%% Rescanned papers and wrote the paragraph below.
%% DO NOT PERMANENTLY DELETE!
%%
% Unfortunately, maintained of several host states at different layers,
% renders any content migration a costly process. Additionally, the method does
% not make any continuing effort to remap
% the overlay structure after nodes successfully join resulting into poor
% adaptation to changing conditions. Although this approach greatly reduces the
% routing latency to far nodes, it is unable to dynamically identify nodes that
% are close to routers and gateways in order to construct the secondary overlay.
% Nevertheless, static recognition of such nodes is currently done based on BGP
% reports and pre-chosen landmarks, sacrificing the self-organizing attribute of
% traditional DHTs.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Unable to identify nodes that are close to routers but it can reduce routing
% latency to far away node by approximately 50-75%.
medium &
% maintenance of node states, subscriptions and notifications need global
% overview of the overlay resulting in high control overhead
high &
% Statically implemented sacrificing the decentralized nature of DHTs and
% thus poorly scaling. Also we can end up with a very large number of regions.
% Because of the fact that a peer may appear in multiple
% such maps we can conclude that additional traffic for probing and notifications
% renders the approach does not scalable.
low
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Mithos}
\emph{Mithos} \cite{WR2003} is a \p\ protocol which incorporates directed
incremental probing to find near optimal node placement, and is an integration
of geographic layout and proximity routing overlay optimization methods.

The bootstrap phase of \emph{Mithos} starts with a subset of existing members as
the first set of candidates, and while the closest nodes are detected by iteratively
probing the neighborhood, the candidate neighbor list is updated. To
avoid local minima, \emph{Mithos} probes all the neighbors within a two-hop distance
from the current minimum before concluding the process.

After finding its first neighbor, the new-coming peer is assigned an ID using
information gathered during the iterative neighbor selection phase. Based on the
two closest nodes found, along with these nodes neighbors and the corresponding
distances a system of virtual springs is set so that the tension of each spring
is inversely proportional to the distance. The system is then left to settle,
and the state in which it settles is used for its ID. The benefit of these
synthetic coordinates are that
they are explicitly used as the node's ID, and distance computations among nodes
can be done in ID space without requiring physical measurements.

The last step of the algorithm is the interconnection among neighbors. \emph{Mithos}
uses a \emph{quadrant-based} mechanism in which each node establishes
a link to the closest neighbor in each quadrant. During forwarding, the next
hop is performed towards a neighbor in the same quadrant as the final
destination. The new coming node may not know of other neighbors in all
quadrants, therefore, the node first identifies neighbors in all quadrants
using a mechanism based on ideas similar to a perimeter walk\footnote{Used in
the Greedy Perimeter Stateless Routing (GPSR) protocol.} and then improves the
results using parallel path processing by taking into account further geometric
properties of node relationships.

One major limitation of \emph{Mithos} is that the protocol cannot efficiently handle
dynamic arrivals and departures from the network. Moreover, parts of this spring
relaxation technique requires a centralized
implementation~\cite{cox_vivaldi_2004}.

%% VM I have read this somewhere but I didn't keep the reference ... :(
%%
% In order to avoid local minima during neighbor detection, extensive probing
% must be undertaken. In simulation, unfortunately, only very small-sized
% overlay topologies (of 200 to 1000 nodes) have been used and thus no safe
% conclusions can be made as for the behavior of an extensively large,
% real-world \p\ deployment of the scheme.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{LAPTOP}
\emph{LAPTOP} \cite{WLH2007}, organizes the overlay into a tree-based hierarchy
with the goals of reducing hops during message routing as well as minimizing
maintenance overhead. Additionally, a caching scheme is also incorporated so as
to further reduce routing table update costs. It is theoretically shown
that in \emph{LAPTOP}, routing path length is bounded by $O(log_d N)$ and node joining
and leaving in the overlay network is bounded by $O\left( d log_d N \right)$
hops in a balanced overlay tree, where $N$ is the number of nodes, and $d$ is
the maximum degree of each node. \emph{LAPTOP} implements the geographical layout
approach  and constructs its layout in a self-organizing and efficient fashion.
The physical distances are estimated with round trip times to a few existing
nodes in the overlay network.

@@

The routing scheme is similar to the longest-prefix IP matching scheme. At each
forwarding hop, any message travels up the tree until the first common ancestor
of both source and destination nodes is reached and then starts descending to
arrive to its target. During tree traversals, special entries in the routing
tables, called \emph{routing cache}, are maintained to increase routing
efficiency and achieve finer load balance. Caching enables a node to forward a
message to a better longest-prefix match than that of its direct ancestor making
a larger, quicker and more cost effective step through the overlay and toward
the destination. To improve scalability, the number of child nodes and the
size of the routing cache are limited. In terms of overlay maintenance, \emph{LAPTOP}
incorporates a simple \emph{heartbeat}-based technique where each parent node is
responsible for monitoring its children.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{Proximity in Kademlia}
% Before discussing {\em Proximity in Kademlia}, it would be helpful to briefly
% summarize the highlights about the {\em Kademlia} protocol. {\em Kademlia}
% \cite{maymounkov_kademlia_2002} is a distributed hash table (DHT) based
% protocol for \p\ networks, which uses an iteration based lookup algorithm. The
% protocol uses the standard $160$ bit ID system for nodes and locates the nodes
% in a prefix binary tree, where IDs are used as prefixes. The iterative lookup
% operations are done over this prefix binary tree, which converges to
%logarithmic
% lookup times. The ID's are assigned randomly, therefore in Kademlia, there is
%no
% proximity control, which results in inefficient use of the underlying network
% during lookup and retrieve operations.

\cite{KLKP2008} introduces proximity
controls over the vanilla Kademlia protocol's random ID assignment to
optimize the underlying network usage.
% The work in this paper
% can also be considered as a more generic study of incorporating proximity into
% iterative lookup algorithms.
The overlay tier improves its routing performance using a cost metric that is
provided by the underlying tier. Such information can be acquired by
\begin{inparaenum}[\itshape i\upshape)]
  \item measuring previous lookups,
  \item exchanging measurements of other peers or jointly calculating them with
  others, or
  \item using a local database to look up information.
\end{inparaenum}

Regarding routing optimization both techniques of \emph{proximity routing
selection (PRS)} and \emph{proximity neighbor selection (PNS)} are studied.
%%
In \emph{PRS}, the goal is to choose the best next hop during routing a message.
% Due to the fact that the routing in Kademlia is iterative, it is the initiator
% of the lookup that chooses each next hop from a set of candidate nodes.
Plain old Kademlia (\emph{POK}) always chooses the closest node with respect to
the XOR metric but with proximity routing, Kademlia should also choose the node
that exhibits the smallest underlay metric cost, balancing a trade-off between
the overlay and underlay progress distances.
%%
In \emph{PNS}, on the other hand, the goal is to keep peers with the least
contact cost in the routing table. As Kademlia continuously learns about new
peers via incoming requests and an iterative lookup process, no special
algorithm for searching more cost-effective peers is necessary and it simply
chooses the best peers seen so far.

Proximity injection in Kademlia improves locality of connections between peers.
Towards the enhancement of traffic locality, it is suggested to use information
from the MaxMind GeoIP database\footnote{http://www.maxmind.com} and use it to
form clusters of nearby peers. An other alternative method could use Vivaldi
coordinates \cite{cox_vivaldi_2004}, however, it is suggested that the
clustering approach performs better. Specifically when combined with both
\emph{PRS} and \emph{PNS} preserves more than $40\%$ of the messages within an
ISP and allows only a small $6\%$ to go through intercontinental backbones.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Uses static information to cluster nodes which is not the most efficient
% approach in dynamic systems
% PNS reduces lookup latencies by 66\% compared to POK
medium &
% 
medium &
% MaxMind is commercial and cannot be used by FOSS P2P systems. Trying to
% expose information from P2P client can result in security issues.
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{CHOord considering Proximity on IPv6 (CHOP6)}
\emph{CHOP6} \cite{MT2007} roughly estimates the proximity among
nodes by exploiting the IPv6 address format and RTT information, if available.
The proximity estimation is achieved by introducing a 64-bit ID scheme in which
the least significant bit part is the IPv6 global routing prefix and thus
enabling a longest prefix match scheme. The protocol is designed based on the
observation that it is possible to estimate a node's geographical location by
simply examining the upper 32-bits of its IPv6 address. Moreover, similar to
Chord\cite{stoica_chord_2001}, CHOP6 uses a finger table, whose entries hold
more than one candidate node.

@@

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Simple approach that exploits proximity but it is coarse grained and thus not
% efficient.
low &
% Its static nature imposes low overhead to the network
low &
% IPv6 is still not widely applied and supported.
medium
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{T2MC}
\emph{T2MC} \cite{SLCGZ2008} uses traceroute logs to detect the AS boundaries and
cluster nearby peers together to reduce the redundant multiple message
passes between the AS boundaries. \emph{T2MC} uses a customized k-mean classification
algorithm with $k = 2$ to perform the classification, and exploits the stable
structure of the Internet routers to guide clustering. Even though trace-route
provides detailed information about the network structure, use of trace-route
creates overhead on the overall network structure. For this same reason, it is
not unusual for network administrators to disable this support on their network
routers, which may affect the performance of the \emph{T2MC} algorithm.

The results of T2MC are encouraging but it is also considered coarse-grained
since it classifies peers into only ``remote'' and ``close''
groups \cite{QLZG2009}.

%T2MC \cite{SLCGZ2008} uses trace-route logs to detect different autonomous
%systems and cluster close by peers, thus reducing redundant traffic that crosses
%autonomous system boundaries. T2MC uses a customized k-mean classification
%algorithm with $k=2$ to perform the classification, and exploits the stable
%structure of the Internet routers to guide clustering. The peer chooses the
%minimum and maximum Latency results from the trace-route to initialize the
%centroids of two sets. Then the peer calculates, for all hops along its
%trace-routed path, the peer calculates the absolute distance to these centroids
%of both sets and assigns the routers to that centroid with which it has the
%smaller absolute distance. After that, the mean and variance values are
%calculated and if the variance is larger than a predefined threshold then the
%algorithm sets the two latency mean values as new centroids for the two sets and
%loops again. Ultimately peer will end up with two sets having the minimum
%intra-set variance. The peer, then, chooses the router from the second set with
%the minimum hops attribute and sets it as a threshold. The selected router and
%all others whose hops attribute is larger than the threshold are classified as
%\emph{remote} router cluster. The remaining are classified as \emph{near}. From
%the near class, the peer chooses the one with maximum hops attribute as its edge
%router, and registers it along with all near clusters into the DHT of the
%\p\ overlay. As new peers join the network, those that share the same edge
%router or any of the members of the near router clusters, would form a close
%peer cluster. As Edge routers can provide more valuable information than other
%members of the near set, T2MC was designed to prioritize interaction of peers
%and edge gateways.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{PChord}
\emph{PChord} \cite{HLYW2005} is based on the Chord DHT and adds
proximity into its routing mechanism. The main modification over the
vanilla Chord, is the inclusion of a \emph{proximity list} into its routing table
so that the next hop is decided based on both the best progress
towards the key and the physical proximity of the candidate nodes.
The list is, at join time, empty but as the \emph{PChord} node starts to interact with
other nodes it applies a heuristic mechanism to fill the list. Entries are
dynamically added or removed as the network state changes. The overlay
maintenance cost is similar to Chord's. Apart from the costs incurred by node
join/leave and by heart-beat messages for connection verification, \emph{PChord} adds
heart-beat among nodes inside the proximity list. Fortunately, since the list
contains nodes with the lowest RTT, the added cost is kept to a minimum.

Due to the fact that, with each step, there is always progress towards the
target node in the ID space, \emph{PChord} will result in a reduced hop number than
Chord, since each hop in \emph{PChord} is larger or at least equal in the key space.
Passing through proximity links, in the underlying network, means reduced routing
cost. Additionally, if the number of entries in the proximity list is the same
as the number of network partitions, \emph{PChord} prevents hops from jumping in and
out of the network partition the current node belongs to.

\emph{PChord}’s advantages include the fact that the main routing optimizations are of
less overlay hops, following proximity links and crossing the same network
partition only once along a routing path while in the same time keeping Chord's
light-weight maintenance cost. On the other hand, it exhibits slow convergence
speed and inefficiency in a high churn environment\cite{DK2006}.

\begin{center}
\begin{tabular}{ccc}
\emph{Efficiency} & \emph{Overhead} & \emph{Scalability} \\
\hline
% Small decrease in needed overlay hops per object publishing but rigid in
% high churn environments.
medium &
% The are heartbeat mechanism for both finger table and proximity list adds a
% burden to the network by default. But the overall overhead is decreased as
% the proximity list becomes longer.
medium &
% Improvement of routing is achieved with more object pointers helping to
% shorten the routing path of object querying. When the distance of query source
% to target is far away both Chord and PChord achieve the average hop number to
% not increase.
high
\end{tabular}
\end{center}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \subsubsection{AChord}
\emph{AChord} \cite{DK2006} uses IPv6's any-cast mechanism to
relieve the protocol from complex joining procedures, and achieve high accuracy
of network proximity and thus, high routing efficiency. Any-cast delivers a message
coming from the outside of an any-cast group to the physically closest node in
that group \cite{M2002}. \emph{AChord} organizes all nodes participating in the
overlay network into an any-cast group. Any joining node comes outside of that group
and, thus, is
automatically forwarded to the physically nearest node in order to bootstrap,
avoiding the need to, explicitly, maintain such nodes, and the effort of finding
a way of locating the physically nearest among them. The ID of the new-coming
node is computed based on the ID of the bootstrap node and the bootstrap's
predecessor in a way that the constructed ID will be positioned between the
aforementioned two. After joining, the finger table is built the same way as in
the Chord protocol. Moreover, additional nodes are maintained in a structure
called \emph{adjacency table} which stores information about the closest known
node. The routing decision is made by using both the adjacency and finger
tables.
